### 2018-03-09 paper/background.tex
Add draft of S3
Expand engineering aspects section

\Make\footnote{There are numerous implementations of \Make and none comes with a
formal specification. In this paper we therefore use a simple and sensible
approximation to a real \Make that you might find on your machine.} was developed
more than 40 years ago to automatically build software libraries and executable
programs from source code. It uses \emph{makefiles} to describe build tasks and
their dependencies in a simple text form. For example:

@@

Note that \Excel's approach to marking cells as dirty is a relatively minor
variation of the approach used by \Make: a cell is considered dirty if it was
modified since the previous build \emph{or if it is volatile}, where the notion
of volatility includes dynamic dependencies, as well as \emph{non-determinism}
that will be discussed in~\S\ref{sec-engineering}. We therefore put both \Make
and \Excel in the same category of build systems that associate a \emph{dirty bit}
with every file or cell. (To avoid the ``file or cell'' nuisance, we introduce the
abstract notion of \emph{keys} in~\S\ref{sec-background-vocabulary}.)

@@

To support cloud builds, \Bazel maintains a conventional \emph{content-addressable
cache}, which can be used to fetch a previously built file given the hash of its
content, as well as \emph{dependency graphs from all previous builds}, annotated
with observed file hashes\footnote{Here we ignore the issue of limited cloud
storage resources for the sake of simplicity; in practice, old entries are
regularly evicted from the storage, as further discussed in~\S\ref{sec-engineering}.}.
The latter allows to bypass the execution of a task, by predicting the hash of
its result from the hashes of the dependencies, and subsequently fetch the result
from the cache. A concrete implementation is provided in~\S\ref{sec-examples}.

>>>>

\Make\footnote{There are numerous implementations of \Make and none comes with a
formal specification. In this paper we therefore use a simple and sensible
approximation to a real \Make that you might find on your machine.} was developed
more than 40 years ago to automatically build software libraries and executable
programs from source code. It uses \emph{makefiles} to describe tasks (often
referred to as \emph{build rules}) and their dependencies in a simple text form.
For example:

@@

Note that \Excel's approach to marking cells as dirty is a relatively minor
variation of the approach used by \Make: a cell is considered dirty if it was
modified since the previous build \emph{or if it is volatile}, where the notion
of volatility includes dynamic dependencies, as well as \emph{non-determinism}
that will be discussed in~\S\ref{sec-non-determinism}. We therefore put both
\Make and \Excel in the same category of build systems that associate a
\emph{dirty bit} with every file or cell. (To avoid the ``file or cell''
nuisance, we introduce the abstract notion of \emph{keys}
in~\S\ref{sec-background-vocabulary}.)

@@

To support cloud builds, \Bazel maintains a conventional \emph{content-addressable
cache}, which can be used to fetch a previously built file given the hash of its
content, as well as \emph{dependency graphs from all previous builds}, annotated
with observed file hashes\footnote{Here we ignore the issue of limited cloud
storage resources for the sake of simplicity; in practice, old entries are
regularly evicted from the storage, as further discussed
in~\S\ref{sec-sec-cloud-aspects}.}. The latter allows to bypass the execution of
a task, by predicting the hash of its result from the hashes of the dependencies,
and subsequently fetch the result from the cache. A concrete implementation is
provided in~\S\ref{sec-examples}.

### 2018-03-09 paper/engineering.tex
Expand engineering aspects section

\begin{itemize}
    \item Sandboxing
    \item Non-determinism, specifically, its useless form. This doesn't
          include \Excel's \textsf{RANDBETWEEN} function, which is useful.
          To define correctness of non-deterministic build systems, one needs to
          switch from \hs{Applicative} and \hs{Monad} abstractions to
          \hs{Alternative} and \hs{MonadPlus}, respectively, but this is not
          required for actual implementations, most of which can happily accept
          non-deterministic results (with a notable exception of \Buck).
    \item Concurrency: nothing fundamentally difficult and fits our theory, but
          hard to get right in practice.
    \item Cloud aspects: eviction/garbage collection, `frankenbuilds', etc.
    \item Iterative computations, e.g. LaTeX.
    \item Polymorphic keys and values: tasks with multiple outputs, e.g. GHC's
          \cmd{hi} and \cmd{o} files, phony tasks, etc.
\end{itemize}

\subsection{Non-deterministic computations}

@@

>>>>

\subsection{Non-deterministic computations}\label{sec-non-determinism}

Sandboxing: guarding against non-determinism due to missing dependencies.

\todo{AM}{Rewrite everything below.}

Non-determinism, specifically, its \emph{useless} form. This doesn't
include \Excel's \textsf{RANDBETWEEN} function, which is useful.
To define correctness of non-deterministic build systems, one needs to
switch from \hs{Applicative} and \hs{Monad} abstractions to
\hs{Alternative} and \hs{MonadPlus}, respectively, but this is not
required for actual implementations, most of which can happily accept
non-deterministic results (with a notable exception of \Buck).

... Some old text:

@@

\subsection{Concurrency}\label{sec-concurrency}

Nothing fundamentally difficult and fits our theory, but hard to get right in
practice.

\todo{NM}{Add a paragraph on how to introduce concurrency to implementations
in~\S\ref{sec-examples}}.

\subsection{Cloud aspects}\label{sec-cloud-aspects}

Cloud aspects: eviction/garbage collection, `frankenbuilds', etc.

\todo{AM}{Descibe frankenbuilds linking to CloudBuild paper.}

\subsection{Tracking and self-tracking}\label{sec-tracking-aspects}

Annotating dependency graphs with hashes, times, etc. is in many cases just an
engineering trade-off.

\todo{NM}{Clarify, e.g. Shake's \cmd{ChangeModtime}.}

\todo{NM}{Explain how self-tracking can be implemented.}

\subsection{Iterative computations}\label{sec-iterative-compute}

Iterative computations, e.g. LaTeX. \Excel has built-in support for this.

\todo{NM}{Clarify how this fits with the example implementations.}

\subsection{Polymorphism}\label{sec-polymorphism}

Polymorphic keys and values: tasks with multiple outputs, e.g. GHC's \cmd{hi}
and \cmd{o} files, phony tasks, etc.

\todo{AM}{Give an example.}

\todo{NM}{Describe how to add this feature to \Shake.}
